Ordinal regression models made easy. A tutorial on parameter interpretation, data simulation, and power analysis.

### 2023-12-12 paper/paper-new.tex

\begin{verbatim}
## 
## Call:
## lm(formula = ys ~ x1 + x2, data = dat)
## 
## Residuals:
##     Min      1Q  Median      3Q     Max 
## -4.5222 -0.6765  0.0000  0.6741  4.3425 
## 
## Coefficients:
##             Estimate Std. Error t value Pr(>|t|)    
## (Intercept) 0.001086   0.003159   0.344    0.731    
## x1          0.998374   0.003167 315.209   <2e-16 ***
## x2          0.502929   0.003163 159.004   <2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 0.9989 on 99997 degrees of freedom
## Multiple R-squared:  0.5549, Adjusted R-squared:  0.5549 
## F-statistic: 6.233e+04 on 2 and 99997 DF,  p-value: < 2.2e-16
\end{verbatim}

\begin{verbatim}
## formula: y ~ x1 + x2
## data:    dat
## 
##  link   threshold nobs  logLik     AIC       niter max.grad cond.H 
##  probit flexible  1e+05 -122644.47 245300.95 5(0)  4.75e-08 9.0e+00
## 
## Coefficients:
##     x1     x2 
## 0.9974 0.5013 
## 
## Threshold coefficients:
##     1|2     2|3     3|4     4|5 
## -0.8473 -0.2529  0.2551  0.8449
\end{verbatim}

\begin{verbatim}
## (Intercept)          x1          x2 
## 0.001086866 0.999502295 0.503497031
\end{verbatim}

@@

\begin{verbatim}
## $sigma
## [1] 1.79423
## 
## $b0
## [1] 2.999998
## 
## $beta
## [1] 1.786346
## 
## $alpha
## [1] 1.500000 2.543349 3.447639 4.500000
\end{verbatim}

@@

\begin{verbatim}
##            1         2         3         4         5
## 1 0.19881313 0.1989910 0.1974955 0.1998745 0.2048259
## 2 0.07912516 0.1070303 0.1513017 0.2359672 0.4265757
\end{verbatim}

@@

\begin{verbatim}
##         xa         xb 
## -0.9791059 -0.9791059
\end{verbatim}

@@

\begin{verbatim}
## Calls:
## 1: glm(formula = y123vs4 ~ x, family = binomial(link = "logit"), data = dat)
## 2: glm(formula = y12vs34 ~ x, family = binomial(link = "logit"), data = dat)
## 3: glm(formula = y1vs234 ~ x, family = binomial(link = "logit"), data = dat)
## 
##               Model 1   Model 2   Model 3
## (Intercept)  1.098911 -0.000607 -1.096559
## SE           0.015143  0.016352  0.021668
##                                          
## x             -5.0004   -4.9958   -5.0132
## SE             0.0364    0.0484    0.0736
## 
\end{verbatim}

@@

\begin{verbatim}
##    p_y1  p_y2  p_y3  p_y4
## 1  0.57 0.232 0.134 0.063
## 2 0.128 0.194 0.262 0.415
## 3   0.1 0.172 0.255 0.472
## 4 0.043 0.105 0.207 0.645
## 5   ...   ...   ...   ...
## 6 0.381 0.264 0.207 0.147
## 7 0.459 0.257 0.177 0.106
## 8 0.332 0.263 0.225  0.18
## 9 0.439  0.26 0.185 0.116
\end{verbatim}

@@

@@

{\centering \includegraphics{paper-new_files/figure-latex/unnamed-chunk-18-1}

@@

\caption{ }\label{fig:unnamed-chunk-18}
\end{figure}

@@

  \CommentTok{\# get all the correct functions according to the link}
\NormalTok{  lf }\OtherTok{\textless{}{-}} \FunctionTok{get\_link}\NormalTok{(link)}
  \ControlFlowTok{if}\NormalTok{(}\FunctionTok{is.null}\NormalTok{(alphas))\{}
    \CommentTok{\# calculate thresholds if not provided}
\NormalTok{    alphas }\OtherTok{\textless{}{-}} \FunctionTok{probs\_to\_th}\NormalTok{(probs, }\AttributeTok{link =}\NormalTok{ link)}
\NormalTok{  \}}
\NormalTok{  k }\OtherTok{\textless{}{-}} \FunctionTok{length}\NormalTok{(alphas) }\SpecialCharTok{+} \DecValTok{1} \CommentTok{\# number of ordinal outcomes}
\NormalTok{  n }\OtherTok{\textless{}{-}} \FunctionTok{nrow}\NormalTok{(data) }\CommentTok{\# number of observations}

@@

  \ControlFlowTok{if}\NormalTok{(}\FunctionTok{is.null}\NormalTok{(th))\{}
\NormalTok{    th }\OtherTok{\textless{}{-}} \FunctionTok{probs\_to\_th}\NormalTok{(probs, link)}
\NormalTok{  \}}

@@

\NormalTok{    th }\OtherTok{\textless{}{-}} \FunctionTok{probs\_to\_th}\NormalTok{(probs, link)}

>>>>

\begin{verbatim}
## 
## Call:
## lm(formula = ys ~ x1 + x2, data = dat)
## 
## Residuals:
##     Min      1Q  Median      3Q     Max 
## -4.0357 -0.6735 -0.0024  0.6761  4.0977 
## 
## Coefficients:
##              Estimate Std. Error t value Pr(>|t|)    
## (Intercept) 0.0004437  0.0031671    0.14    0.889    
## x1          0.9995687  0.0031605  316.27   <2e-16 ***
## x2          0.5046426  0.0031633  159.53   <2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 1.002 on 99997 degrees of freedom
## Multiple R-squared:  0.5577, Adjusted R-squared:  0.5576 
## F-statistic: 6.303e+04 on 2 and 99997 DF,  p-value: < 2.2e-16
\end{verbatim}

\begin{verbatim}
## formula: y ~ x1 + x2
## data:    dat
## 
##  link   threshold nobs  logLik     AIC       niter max.grad cond.H 
##  probit flexible  1e+05 -121998.29 244008.57 5(0)  6.33e-08 9.1e+00
## 
## Coefficients:
##     x1     x2 
## 1.0025 0.5039 
## 
## Threshold coefficients:
##     1|2     2|3     3|4     4|5 
## -0.8418 -0.2514  0.2543  0.8413
\end{verbatim}

\begin{verbatim}
##  (Intercept)           x1           x2 
## 0.0004430131 0.9980388055 0.5038702203
\end{verbatim}

@@

\begin{verbatim}
## $sigma
## [1] 1.776447
## 
## $b0
## [1] 2.999548
## 
## $beta
## [1] 1.777515
## 
## $alpha
## [1] 1.500000 2.549889 3.449757 4.500000
\end{verbatim}

@@

\begin{verbatim}
##            1         2         3         4         5
## 1 0.20301120 0.1911931 0.1969380 0.2054224 0.2034352
## 2 0.07967757 0.1014347 0.1483835 0.2414729 0.4290314
\end{verbatim}

@@

\begin{verbatim}
##         xa         xb 
## -0.9379058 -0.9379058
\end{verbatim}

@@

\begin{verbatim}
## Calls:
## 1: glm(formula = y123vs4 ~ x, family = binomial(link = "logit"), data = dat)
## 2: glm(formula = y12vs34 ~ x, family = binomial(link = "logit"), data = dat)
## 3: glm(formula = y1vs234 ~ x, family = binomial(link = "logit"), data = dat)
## 
##             Model 1 Model 2 Model 3
## (Intercept)  1.0844 -0.0176 -1.1199
## SE           0.0151  0.0163  0.0216
##                                    
## x           -4.9826 -4.9374 -4.9444
## SE           0.0364  0.0481  0.0733
## 
\end{verbatim}

@@

\begin{verbatim}
##    p_y1  p_y2  p_y3  p_y4
## 1 0.481 0.254 0.169 0.096
## 2 0.092 0.165 0.252 0.491
## 3 0.098  0.17 0.254 0.478
## 4 0.426 0.261  0.19 0.122
## 5   ...   ...   ...   ...
## 6  0.11  0.18 0.258 0.452
## 7 0.315 0.261 0.231 0.193
## 8 0.108 0.179 0.258 0.456
## 9 0.065 0.136 0.234 0.565
\end{verbatim}

@@

\begin{longtable}{llrrrrr}
\toprule
coef.type & term & estimate & std.error & statistic & p.value & row \\ 
\midrule
\$\$\textbackslash{}alpha\$\$ & 1|2 & -0.680 (*-0.674*) & $0.004$ & $-152.618$ & $0.000$ & $1.000$ \\ 
 & 2|3 & -0.005 (*0.000*) & $0.004$ & $-1.179$ & $0.238$ & $2.000$ \\ 
 & 3|4 & 0.666 (*0.674*) & $0.004$ & $149.925$ & $0.000$ & $3.000$ \\ 
\$\$\textbackslash{}mu\$\$ & x & 0.495 (*0.500*) & $0.004$ & $132.696$ & $0.000$ & $1.000$ \\ 
\bottomrule
\end{longtable}

@@

{\centering \includegraphics{paper-new_files/figure-latex/unnamed-chunk-17-1}

@@

\caption{ }\label{fig:unnamed-chunk-17}
\end{figure}

\begin{longtable}{llrrrrr}
\toprule
coef.type & term & estimate & std.error & statistic & p.value & row \\ 
\midrule
\$\$\textbackslash{}alpha\$\$ & 1|2 & -0.681 (*-0.674*) & $0.004$ & $-152.568$ & $0.000$ & $1.000$ \\ 
 & 2|3 & -0.007 (*0.000*) & $0.004$ & $-1.708$ & $0.088$ & $2.000$ \\ 
 & 3|4 & 0.664 (*0.674*) & $0.004$ & $149.541$ & $0.000$ & $3.000$ \\ 
\$\$\textbackslash{}mu\$\$ & x & 0.498 (*0.500*) & $0.004$ & $133.483$ & $0.000$ & $1.000$ \\ 
\bottomrule
\end{longtable}

@@

  \CommentTok{\# get all the correct functions according to the link}
\NormalTok{  lf }\OtherTok{\textless{}{-}} \FunctionTok{get\_link}\NormalTok{(link)}
  \ControlFlowTok{if}\NormalTok{(}\FunctionTok{is.null}\NormalTok{(alphas))\{}
    \CommentTok{\# calculate thresholds if not provided}
\NormalTok{    alphas }\OtherTok{\textless{}{-}} \FunctionTok{prob\_to\_alpha}\NormalTok{(probs, }\AttributeTok{link =}\NormalTok{ link)}
\NormalTok{  \}}
\NormalTok{  k }\OtherTok{\textless{}{-}} \FunctionTok{length}\NormalTok{(alphas) }\SpecialCharTok{+} \DecValTok{1} \CommentTok{\# number of ordinal outcomes}
\NormalTok{  n }\OtherTok{\textless{}{-}} \FunctionTok{nrow}\NormalTok{(data) }\CommentTok{\# number of observations}

@@

  \ControlFlowTok{if}\NormalTok{(}\FunctionTok{is.null}\NormalTok{(th))\{}
\NormalTok{    th }\OtherTok{\textless{}{-}} \FunctionTok{prob\_to\_alpha}\NormalTok{(probs, link)}
\NormalTok{  \}}

@@

\NormalTok{    th }\OtherTok{\textless{}{-}} \FunctionTok{prob\_to\_alpha}\NormalTok{(probs, link)}
